Population Ecology
========================================================
```{r setglobalopts, echo = F}
opts_chunk$set(comment = NA, fig.width=7, fig.height=6)
```


```{r simpleEQ}
N.t = 30
R = 1.16

N.t1 <- N.t * R
N.t1
```

We can use this equation for multiple years and plot the result: 

```{r ExpGrowth}
N <- c()
for(i in 1:10){
  N[i] <- N.t * R ^ i
}

plot(N)
```



## Exercise Examples

1.1 Blue Whale Recovery
```{r}
N.1963 <- 10000
N.target <- 50000

R <- 1.10

t <- log(50000/10000) / log(R)
t

R <- 1.02

t <- log(50000/10000) / log(R)
t
```  
  
1.2 Human Population (1800 - 1995)  
```{r}
humans <- data.frame(Year = c(1800, 1850, 1870, 1890, 1910, 1930, 1950, 1970, 1975, 1980, 1985, 1990, 1995))
humans <- cbind(humans, Population = c(0.91, 1.13, 1.30, 1.49, 1.70, 2.02, 2.51, 3.62, 3.97, 4.41, 4.84, 5.29, 5.75))

Time.Interval <- c()
Time.Interval[1] <- NA
for (i in 2:length(humans$Year)){
  Time.Interval[i] <- humans$Year[i] - humans$Year[i-1]
}

Prev.Population <- humans$Population[1:(length(humans$Population)-1)]

Growth.Rate <- humans$Population[2:length(humans$Population)] / Prev.Population

Annual.Growth.Rate <- Growth.Rate ^ (1 / Time.Interval[2:length(Time.Interval)])

Table_1.3 <- cbind(humans, Time.Interval, 
                   Prev.Population = c(NA, Prev.Population), 
                   Growth.Rate = c(NA, Growth.Rate), 
                   Annual.Growth.Rate = c(NA, Annual.Growth.Rate))
Table_1.3
```  

```{r echo = FALSE, include = F}
require(ggplot2)
```

Plot using Growth Rate  

```{r}
qplot(x = Table_1.3$Year[2:13], y = Table_1.3$Growth.Rate[2:13], 
      xlab = "Year", ylab = "Growth Rate")
```
  
Plot using the Annual Growth Rate  

```{r}
qplot(x = Table_1.3$Year[2:13], y = Table_1.3$Annual.Growth.Rate[2:13], 
      xlab = "Year", ylab = "Annual Growth Rate")
```

Compare the change in annual growth rate with the absolute increase in number of people

```{r}
Table_1.4 <- data.frame(Year = c(1975, 1985, 1995), Population.Size = c(3.97, 4.84, 5.75),
                        Annual.Growth.Rate = c(1.018630, 1.018782, 1.016816))
Number.Added <- Table_1.4$Population.Size * (Table_1.4$Annual.Growth.Rate - 1)

Table_1.4 <- cbind(Table_1.4, Number.Added)

Table_1.4
```

Calculate (based on the number added in 1995) the number added:   
 * per day:  
```{r}
per.day <- Table_1.4$Number.Added[3] / 365
per.day
```  
 * per hour:  
```{r}
per.hour <- per.day / 24
per.hour
```  
 * per minute:
```{r}
per.minute = per.hour / 60
per.minute
```  
  
1.3 Human Population (1995 - 2035)

```{r}
Table_1.5 <- data.frame(Year = c(1995, 2005, 2015, 2025, 2035))

fec <- 0.0273
R <- 1.016816
s <- R - fec

change10yr <- (R - 1) / 4

R05 <- R - change10yr
R15 <- R05 - change10yr
R25 <- R15 - change10yr

Rs <- c(R, R05, R15, R25, 1)

fecs <- Rs - s

Rs10 <- Rs ^ 10

Population <- 5.75
for(i in 2:5){
  Population[i] <- Population[i-1] * Rs10[i-1]
}


Table_1.5 <- cbind(Table_1.5, Fecundity = fecs, R = Rs, R_10_yr = Rs10, Population = Population)
Table_1.5
```

